Modeling believable lighting is a crucial component of computer graphics applications, including games and modeling programs. Physically accurate lighting is complex and is not currently feasible to compute in real-time situations. Therefore, much research is focused on investigating efficient ways to approximate light behavior within these real-time constraints.

In this thesis, we implement a general purpose algorithm for real-time applications to approximate indirect lighting. Based on voxel cone tracing~\cite{crassin2011interactive}, we use a filtered representation of a scene to efficiently sample ambient light at each point in the scene. We present a novel approach to scene voxelization using hardware tessellation and compare it with an approach utilizing hardware rasterization. We also investigate possible methods of nonuniform voxelization.

Our contributions include a complete and open-source implementation of voxel cone tracing along with both voxelization algorithms. We find similar performance and quality with both voxelization algorithms.

%We also extend the base algorithm to better accomodate large scenes by applying a warping function to our filtered representation. This allows us to increase lighting detail near the camera without a significant increase in memory usage or a performance penalty.

%Final results show real-time performance and good lighting quality largely independent of screen and scene size. In addition, the implementation is open-source, cross-platform, and easy to integrate into existing engines.